---
title: "GSC Robustness Check: Microsoft AI Applicability Score — COVID-Inclusive Analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

**Robustness variant:** Includes OEWS years 2020 and 2021 (COVID years) in the estimation sample.

Baseline (`gsc_ms_score (3).Rmd`) excluded 2020–2021. This version includes them to test whether COVID volatility substantially changes the estimated treatment effect.

Treatment timing remains: 2022 = final pre-treatment year, 2023 = first post-treatment year (matching DID and SC specifications).

## STEP 1: Load data

```{r load-data}
library(gsynth)
library(dplyr)
library(ggplot2)

df <- read.csv("bls_onet_felten_ms_panel_harmonised.csv")

# COVID-INCLUSIVE: do not exclude years 2020, 2021
# df <- df %>% filter(!(year %in% c(2020, 2021)))  # COMMENTED OUT for COVID-inclusive analysis

# Compute log outcomes
df$log_emp  <- log(df$TOT_EMP)
df$log_wage <- log(df$A_MEDIAN)

# Replace inf with NA (from log(0))
df$log_emp[is.infinite(df$log_emp)]   <- NA
df$log_wage[is.infinite(df$log_wage)] <- NA

cat("Data loaded. Year range:", min(df$year), "to", max(df$year), "\n")
```

## STEP 2: Aggregate to occupation-year level

```{r aggregate}
occ_year <- df %>%
  group_by(OCC_CODE, year) %>%
  summarise(
    OCC_TITLE = first(OCC_TITLE),
    # FIX: sum(x, na.rm=TRUE) returns 0 (not NA) when every element of x is
    # NA -- e.g. an occupation whose TOT_EMP is suppressed/missing across
    # every NAICS industry in a given year. log(0) = -Inf, which is NOT
    # caught by is.na()/is.infinite() checks run earlier on df, since this
    # value is computed fresh here and never re-checked. Guard explicitly.
    log_emp   = {
      total_emp <- sum(TOT_EMP, na.rm = TRUE)
      if (total_emp == 0 || all(is.na(TOT_EMP))) NA_real_ else log(total_emp)
    },
    # Employment-weighted mean of log_wage, excluding missing wage rows
    log_wage  = {
      valid_idx <- !is.na(log_wage) & !is.na(TOT_EMP)
      if (sum(TOT_EMP[valid_idx], na.rm = TRUE) == 0) {
        NA
      } else {
        weighted.mean(log_wage[valid_idx], TOT_EMP[valid_idx], na.rm = TRUE)
      }
    },
    MS_SCORE   = first(MS_SCORE),
    .groups = "drop"
  )

cat("Occupations:", length(unique(occ_year$OCC_CODE)), "\n")
cat("Years:", paste(sort(unique(occ_year$year)), collapse=", "), "\n")
cat("Any -Inf remaining in occ_year$log_emp:", any(is.infinite(occ_year$log_emp)), "\n")
cat("Rows where log_emp is NA (previously would have been -Inf):",
    sum(is.na(occ_year$log_emp)), "\n")
```

## STEP 3: Create binary treatment variable

Treatment = 1 if (occupation is high-MS_SCORE) AND (year >= 2023)
2022 is the last pre-treatment year (event-time k=-1), matching the DID
estimating equation, `gsc_analysis.Rmd`, and the descriptive notebook --
not 2022 itself, since the May 2022 OEWS survey predates the November
2022 ChatGPT launch entirely.

```{r treatment}
occ_aioe <- occ_year %>%
  group_by(OCC_CODE) %>%
  summarise(MS_SCORE = first(MS_SCORE))

median_aioe <- median(occ_aioe$MS_SCORE, na.rm = TRUE)
cat("AIOE median:", median_aioe, "\n")

high_aioe_occs <- occ_aioe$OCC_CODE[occ_aioe$MS_SCORE >= median_aioe]
cat("High AIOE occupations:", length(high_aioe_occs), "\n")

occ_year$D <- as.integer(
  (occ_year$OCC_CODE %in% high_aioe_occs) & (occ_year$year >= 2023)
)

# Drop rows with missing outcomes or AIOE
occ_year <- occ_year %>% filter(!is.na(MS_SCORE), !is.na(log_emp))

# Convert OCC_CODE to numeric ID (gsynth needs numeric unit index)
occ_year$unit_id <- as.numeric(factor(occ_year$OCC_CODE))

cat("Final panel rows:", nrow(occ_year), "\n")
cat("Unique units:", length(unique(occ_year$unit_id)), "\n")
```

## STEP 4: Check panel balance

gsynth requires a balanced panel.

```{r balance-check}
# Computed from the data rather than hardcoded -- previously fixed at 9,
# which only matched the old 2015-2019+2022-2025 range. With the panel now
# extended back to 2012, a hardcoded 9 would mark every unit "incomplete"
# and silently produce an empty balanced panel.
n_years_expected <- length(unique(occ_year$year))
cat("Expected years per balanced unit:", n_years_expected, "\n")

balance_check <- occ_year %>%
  group_by(unit_id) %>%
  summarise(n_years = n())
cat("Units with complete years:", sum(balance_check$n_years == n_years_expected), "\n")
cat("Units with incomplete years:", sum(balance_check$n_years != n_years_expected), "\n")

# Keep only balanced units
balanced_units <- balance_check$unit_id[balance_check$n_years == n_years_expected]
occ_year_balanced <- occ_year %>% filter(unit_id %in% balanced_units)
cat("Final balanced panel rows:", nrow(occ_year_balanced), "\n")
```

## STEP 5: Run GSC for Employment

```{r gsc-employment}
gsc_emp <- gsynth(
  log_emp ~ D,
  data = occ_year_balanced,
  index = c("unit_id", "year"),
  force = "two-way",
  CV = TRUE,
  r = c(0, 5),
  se = TRUE,
  inference = "parametric",
  nboots = 200,
  parallel = TRUE,
  seed = 42
)

print(gsc_emp)
```

```{r gsc-employment-plot, fig.width=10, fig.height=6}
plot(gsc_emp, type = "gap", main = "GSC: Effect of AI Exposure on Log Employment - MS (COVID-INCLUSIVE)")
```

```{r gsc-employment-counterfactual, fig.width=10, fig.height=6}
plot(gsc_emp, type = "counterfactual", main = "GSC Counterfactual: Log Employment - MS (COVID-INCLUSIVE)")
```

## STEP 6: Run GSC for Wage

```{r gsc-wage}
# gsynth requires a balanced panel. Simply dropping rows with missing
# log_wage (as before) does not guarantee balance -- an occupation missing
# wage in just one year would be left with a gap instead of being dropped
# entirely. Match gsc_analysis.Rmd: keep only occupations with wage data in
# every year of the balanced panel.
wage_complete_units <- occ_year_balanced %>%
  group_by(unit_id) %>%
  summarise(n_valid_wage = sum(!is.na(log_wage))) %>%
  filter(n_valid_wage == n_years_expected) %>%
  pull(unit_id)

cat("Units with complete wage data:", length(wage_complete_units), "\n")

occ_year_balanced_wage <- occ_year_balanced %>%
  filter(unit_id %in% wage_complete_units)

cat("Wage balanced panel rows:", nrow(occ_year_balanced_wage), "\n")
cat("Wage balanced units:", length(unique(occ_year_balanced_wage$unit_id)), "\n")

# Sanity check: both treatment arms must be present, or gsynth/fect will
# error out with "no never-treated units found".
d_by_unit <- occ_year_balanced_wage %>% group_by(unit_id) %>% summarise(ever_treated = max(D))
cat("Never-treated (control) units:", sum(d_by_unit$ever_treated == 0), "\n")
cat("Ever-treated units:", sum(d_by_unit$ever_treated == 1), "\n")

gsc_wage <- gsynth(
  log_wage ~ D,
  data = occ_year_balanced_wage,
  index = c("unit_id", "year"),
  force = "two-way",
  CV = TRUE,
  r = c(0, 5),
  se = TRUE,
  inference = "parametric",
  nboots = 200,
  parallel = TRUE,
  seed = 42
)

print(gsc_wage)
```

```{r gsc-wage-plot, fig.width=10, fig.height=6}
plot(gsc_wage, type = "gap", main = "GSC: Effect of AI Exposure on Log Wage - MS (COVID-INCLUSIVE)")
```

```{r gsc-wage-counterfactual, fig.width=10, fig.height=6}
plot(gsc_wage, type = "counterfactual", main = "GSC Counterfactual: Log Wage - MS (COVID-INCLUSIVE)")
```

## STEP 7: Save results summary

```{r save-summary}
sink("gsc_ms_results_summary_covid.txt")
cat("=== GSC Employment Results (COVID-INCLUSIVE) ===\n")
print(gsc_emp)
cat("\n\n=== GSC Wage Results (COVID-INCLUSIVE) ===\n")
print(gsc_wage)
sink()

cat("Done. Check gsc_ms_results_summary_covid.txt\n")
```
