---
title: "Rambachan-Roth (2023) Sensitivity Analysis -- All Four Specifications"
output: html_document
---

This document runs the Rambachan & Roth (2023) relative-magnitude sensitivity
analysis on the four event-study specifications from Section 4.2 / Eq. (1):
LM_AIOE x {log employment, log wage} and MS_SCORE x {log employment, log wage}.

For each specification it computes:

1. The standard (M-bar = 0) confidence interval, as a benchmark
2. Robust confidence intervals across a grid of relative-magnitude values M-bar
3. The **breakdown value**: the smallest M-bar at which the confidence set
   first includes zero (i.e., the point at which the significant effect
   would no longer be supported)
4. A sensitivity plot for each specification
5. A single summary table across all four specifications

## Step 0: Install and load required packages (run once)

```{r setup, message=FALSE}
required_pkgs <- c("remotes", "dplyr", "ggplot2", "purrr")
for (p in required_pkgs) {
  if (!requireNamespace(p, quietly = TRUE)) install.packages(p)
}

if (!requireNamespace("HonestDiD", quietly = TRUE)) {
  remotes::install_github("asheshrambachan/HonestDiD")
}

library(HonestDiD)
library(dplyr)
library(ggplot2)
library(purrr)
```

## Step 1: Load event-study coefficients and covariance matrices

Inputs are the four pairs of CSV files exported from the Python event-study
script (`did_unified.ipynb`, via the `export_for_honestdid()` helper):

- `coefs_<index>_<outcome>.csv`  -- one column of coefficients, ordered
  chronologically (pre-period years first, then post-period years), with
  the reference year (2022) omitted
- `vcov_<index>_<outcome>.csv`   -- the corresponding full covariance matrix
  (not just standard errors), same row/column order as the coefficient vector

All four specifications share the same panel structure: 8 pre-treatment
years (2012-2019, since 2020-2021 were excluded from the main sample) and
3 post-treatment years (2023-2025), confirmed by the Python export step.

```{r step1_config}
# One entry per specification. Edit paths here if your file names differ.
specs <- list(
  list(index = "LM_AIOE",  outcome = "log_emp",  file_tag = "lm_aioe_log_emp",
       numPrePeriods = 8, numPostPeriods = 3),
  list(index = "LM_AIOE",  outcome = "log_wage", file_tag = "lm_aioe_log_wage",
       numPrePeriods = 8, numPostPeriods = 3),
  list(index = "MS_SCORE", outcome = "log_emp",  file_tag = "ms_score_log_emp",
       numPrePeriods = 8, numPostPeriods = 3),
  list(index = "MS_SCORE", outcome = "log_wage", file_tag = "ms_score_log_wage",
       numPrePeriods = 8, numPostPeriods = 3)
)

load_spec_data <- function(spec) {
  betahat <- as.numeric(read.csv(paste0("coefs_", spec$file_tag, ".csv"))[[1]])
  sigma   <- as.matrix(read.csv(paste0("vcov_", spec$file_tag, ".csv"), header = FALSE))

  # Sanity checks: catch a mismatched file before it silently produces
  # a nonsensical result later in the pipeline.
  stopifnot(length(betahat) == spec$numPrePeriods + spec$numPostPeriods)
  stopifnot(nrow(sigma) == length(betahat), ncol(sigma) == length(betahat))

  list(betahat = betahat, sigma = sigma)
}
```

## Step 2: Run the sensitivity analysis for one specification

`l_vec` targets the simple average of the post-treatment coefficients,
matching the "pooled post-treatment estimate" reported in the dissertation.

```{r step2_function}
run_honestdid_spec <- function(spec, Mbar_grid = seq(0, 2, by = 0.1)) {

  dat <- load_spec_data(spec)
  betahat <- dat$betahat
  sigma   <- dat$sigma

  l_vec <- rep(1 / spec$numPostPeriods, spec$numPostPeriods)

  original <- HonestDiD::constructOriginalCS(
    betahat        = betahat,
    sigma          = sigma,
    numPrePeriods  = spec$numPrePeriods,
    numPostPeriods = spec$numPostPeriods,
    l_vec          = l_vec
  )

  robust <- HonestDiD::createSensitivityResults_relativeMagnitudes(
    betahat        = betahat,
    sigma          = sigma,
    numPrePeriods  = spec$numPrePeriods,
    numPostPeriods = spec$numPostPeriods,
    l_vec          = l_vec,
    Mbarvec        = Mbar_grid
  )

  # Breakdown value: the first Mbar at which the robust CI covers zero.
  breakdown_row <- robust %>%
    mutate(covers_zero = (lb <= 0 & ub >= 0)) %>%
    filter(covers_zero) %>%
    slice(1)

  breakdown_mbar <- if (nrow(breakdown_row) == 0) NA_real_ else breakdown_row$Mbar

  list(
    index          = spec$index,
    outcome        = spec$outcome,
    original       = original,
    robust         = robust,
    breakdown_mbar = breakdown_mbar
  )
}
```

## Step 3: Run all four specifications

```{r step3_run_all}
results_all <- map(specs, run_honestdid_spec)
names(results_all) <- map_chr(specs, ~ paste(.x$index, .x$outcome, sep = " x "))
```

## Step 4: Summary table across all four specifications

This is the table to report in Section 5.4 of the dissertation.

```{r step4_summary}
summary_table <- map_dfr(results_all, function(res) {
  tibble(
    Index                = res$index,
    Outcome              = res$outcome,
    Original_ATT         = res$original$beta_hat,
    Original_CI_lower    = res$original$lb,
    Original_CI_upper    = res$original$ub,
    Breakdown_Mbar        = res$breakdown_mbar
  )
})

print(summary_table)
write.csv(summary_table, "rambachan_roth_summary.csv", row.names = FALSE)
```

**Interpretation of `Breakdown_Mbar`:** the estimated post-treatment effect
remains distinguishable from zero as long as post-treatment departures from
parallel trends are no larger than `Breakdown_Mbar` times the departures
already observed in the pre-treatment period. A larger value indicates a
more robust conclusion; `NA` means the confidence set never covered zero
within the M-bar grid tested (grid may need to be extended -- see Step 6).

## Step 5: Sensitivity plots for each specification

```{r step5_plots, results='asis'}
for (name in names(results_all)) {
  res <- results_all[[name]]
  cat("### ", name, "\n\n")
  fig <- HonestDiD::createSensitivityPlot_relativeMagnitudes(res$robust, res$original)
  print(fig)
  ggsave(paste0("rambachan_roth_", gsub(" x ", "_", tolower(name)), ".png"),
         fig, width = 7, height = 5, dpi = 300)
  cat("\n\n")
}
```

## Step 6: If a breakdown value is NA, extend the grid

If any specification shows `NA` in the summary table, re-run just that one
with a wider M-bar grid before concluding the effect is fully robust:

```{r step6_widen_grid, eval=FALSE}
# Example: re-check the LM_AIOE log_wage specification with a wider grid
wide_result <- run_honestdid_spec(specs[[2]], Mbar_grid = seq(0, 5, by = 0.25))
wide_result$breakdown_mbar
```

## Writing template for Section 5.4

Once the table above is final, replace the placeholder in Section 4.2 /
5.4 of the dissertation with a paragraph along these lines (fill in the
actual numbers from `summary_table`):

> Applying the relative-magnitude sensitivity analysis of Rambachan and
> Roth (2023) to the pooled post-treatment estimates, the breakdown value
> under LM_AIOE is **M-bar = [X]** for wages and **M-bar = [X]** for
> employment; under MS_SCORE the corresponding values are **M-bar = [X]**
> and **M-bar = [X]**. [Comment on whether the wage result is more or less
> robust than the employment result, and whether this is consistent with
> the pre-trend evidence discussed in Section 5.1.]

**Notes:**

- The pre-period is not evenly spaced in calendar time (2019 to 2022 skips
  2020-2021), even though it is evenly spaced in the *event-time index*
  used here (each pre-period year still gets one coefficient, one apart in
  the index). `HonestDiD`'s default assumes evenly spaced relative-time
  periods, which matches how the coefficients were indexed here -- but if
  you want to be fully rigorous about the calendar-time gap, consult the
  package documentation on irregularly spaced periods, or note this
  simplification explicitly as a limitation.
- All four specifications use the *baseline* event study (no Job Zone
  control). If you also want a sensitivity check under the Job Zone-adjusted
  specification, export coefficients/vcov from `did_with_controls_unified`
  the same way and add four more entries to the `specs` list above.
